Best Practices for Multiple Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-06-06 paper/manuscript.tex
add iLLE description and some small tweaks

\subsection{Best practices for reporting data }
\label{sec:plot_data}
\tocsubsectioncomment{@OpenFE team/ @Jenke Scheen/ Volunteers to expand on utility/ limitations of common metrics. Also see @Michael Gilson's GitHub issue.}
Following best practices for data generation and their analysis does not mean that data is reported in the optimal way. As a practitioner of alchemical free energy simulations you also should use best practices for reporting and plotting your results. We encourage the following standard set of analyses and ways to represent data.  Some distinctions can be made between statistics that report how accurate and reproducible your results are to a ground-truth experiment, and how useful a method might be in an applied setting. For the case of relative free energy calculations, results might be reported either per-edge, as DDGs, or as the MLE-augmented DG results. Relative results (and outliers in this regime) can be more indictivate of edge-based errors, whereas outliers in DG could be more indicative of a node-based error, such as a forcefield error for particular ligands. For relative free energy results, it is best practice to plot both.

@@

\begin{figure}
  \includegraphics[width=0.95\linewidth]{figures/fig13_analysis_practices/Figure.png}
  \caption{\textbf{An example of recommended practices for graphing alchemical free energy predictions.} This figure shows the relation between predicted and experimentally-determined Gibbs free energy in kcal/mol with standard errors as error bars. The dark and light-orange regions depict the 1- and 2-kcal/mol confidence bounds. Statistical metrics for the data are reported, with 95\% confidence intervals determined by bootstrapping analysis. Extra care should be taken when investigating potential outliers further.}
   \label{fig:scatterplot_analysis}
\end{figure}

@@

An example of a best practice scatter comparison between computed and experimental values is shown in Fig.~\ref{fig:scatterplot_analysis}, highlighting outliers, error bars and confidence intervals. The data for this plot is artificially generated for illustration purposes.

>>>>

\subsection{Best practices for reporting data }
\label{sec:plot_data}
\tocsubsectioncomment{@OpenFE team/ @Jenke Scheen/ Volunteers to expand on utility/ limitations of common metrics. Also see @Michael Gilson's GitHub issue.}
Following best practices for data generation and their analysis does not mean that data is reported in the optimal way. As a practitioner of alchemical free energy simulations you also should use best practices for reporting and plotting your results. We encourage the following standard set of analyses and ways to represent data. Some distinctions can be made between statistics that report how accurate and reproducible your results are to a ground-truth experiment, and how useful a method might be in an applied setting. For the case of relative free energy calculations, results might be reported either per-edge, as DDGs, or as the MLE-augmented DG results. Relative results (and outliers in this regime) can be more indicative of edge-based errors, whereas outliers in DG could be more indicative of a node-based error, such as a forcefield error for particular ligands. For relative free energy results, it is best practice to plot both.

@@

\begin{figure}
  \includegraphics[width=0.95\linewidth]{figures/fig13_analysis_practices/Figure.png}
  \caption{\textbf{An example of recommended practices for graphing alchemical free energy predictions.} This figure shows the relation between predicted and experimentally-determined Gibbs free energy in kcal/mol with standard errors as error bars as a relative experiment (\textbf{A}) and absolute (\textbf{B}). The dark and light-orange regions depict the 1- and 2-kcal/mol confidence bounds. Statistical metrics for the data are reported, with 95\% confidence intervals determined by bootstrapping analysis. Extra care should be taken when investigating potential outliers further. \textbf{C} demonstrates an example of an iLLE plot to be presented in a prospective setting.}
   \label{fig:scatterplot_analysis}
\end{figure}

@@

An example of a best practice scatter comparison between computed and experimental values is shown in Fig.~\ref{fig:scatterplot_analysis}A-C for relative, absolute and LogP reporting, respectively. These highlight outliers, error bars and confidence intervals correctly. The data for these plots are artificially generated for illustration purposes.
